<<<<<<< HEAD
---
title: "team"
author: "cyber-atlas, Sam Shifflett"
date: "April 11, 2019"
output: html_document
---

```{r}
library('ggplot2')
#acc <- read.csv("https://raw.githubusercontent.com/xdaiISU/ds202materials/master/hwlabs/fars2017/accident.csv", stringsAsFactors = FALSE)

acc<- read.csv('accident.csv', stringsAsFactors = FALSE)
class(acc)

#person <- read.csv("https://raw.githubusercontent.com/xdaiISU/ds202materials/master/hwlabs/fars2017/person.csv")

person <- read.csv('person.csv', stringsAsFactors = FALSE)


```


```{r}
class(acc$DAY_WEEK)

acc$DAY_WEEK <- factor(acc[,'DAY_WEEK'])
levels(acc$DAY_WEEK) <- list("Sunday" = 1, 'Monday'=2, 'Tuesday'=3, 'Wednesday'=4, 'Thursday'=5, 'Friday'=6, 'Saturday'=7)

levels(acc$DAY_WEEK)

ggplot(data=acc, aes(x=DAY_WEEK, fill= DAY_WEEK))+
  geom_bar()+
  #geom_bar(aes(y=stat(count)))+
  geom_text(stat='count', aes(label=..count..), vjust = -.5)+
  scale_y_continuous(limits=c(0,6500))+
  theme(axis.text.x = element_text(angle =45, hjust=1))



```


Show the breakdown of the accidents by hour

```{r}
class(acc$HOUR)

acc$HOUR<- factor(acc[,'HOUR'])

levels(acc$HOUR)[levels(acc$HOUR) == 99] <- "unknown"

levels(acc$HOUR)

ggplot(data=acc, aes(x=HOUR, fill= stat(count), color=HOUR))+
         #fill=stat(count)))+
  geom_bar()+
  #geom_bar(aes(y=stat(count)))+
  geom_text(stat='count', aes(label=..count..), angle=90,hjust = -.5)+
  scale_y_continuous(limits=c(0,2500))+
  theme(axis.text.x = element_text(angle =75, hjust=1))
```



```{r}

class(acc$DRUNK_DR)

acc$DRUNK_DR<- factor(acc[,'DRUNK_DR'])

#levels(acc$DAY_WEEK) <- list("Sunday" = 1, 'Monday'=2, 'Tuesday'=3, 'Wednesday'=4, 'Thursday'=5, 'Friday'=6, 'Saturday'=7)

levels(acc$'DRUNK_DR')

ggplot(data=acc, aes(x=DRUNK_DR, fill=DRUNK_DR, color=stat(count)))+
         #fill=stat(count)))+
  geom_bar()+
 # geom_bar(aes(y=frequency(DRUNK_DR)))+
  geom_text(stat='count', aes(label=..count..), angle=0,vjust = -.1)
  #scale_y_continuous(limits=c(0,3000))
  #theme(axis.text.x = element_text(angle =75, hjust=1))
  #facet_wrap(~DRUNK_DR)


```

##Prepare the Data and Packages
```{r}
acc <- read.csv("https://raw.githubusercontent.com/xdaiISU/ds202materials/master/hwlabs/fars2017/accident.csv", stringsAsFactors = FALSE)
library(tidyverse)
library(ggplot2)
```

##Part One:
###______Number of Crashes Each Day______
```{r Number of Crashes Each Day}
ggplot(acc,aes(x=factor(DAY_WEEK),fill=factor(DAY_WEEK))) + geom_bar(stat_count='identity') + ylab('Number of Crashes') + xlab("Day of the Week") + scale_x_discrete(breaks=c(1,2,3,4,5,6,7),labels=c("Sunday","Monday","Tuesday","Wednesday","Thursday","Friday","Saturday")) + ggtitle("Number of Crashes Each Day") + guides(fill=guide_legend("Days of the Week"))
```

###______Number of Crashes Each Hour______
```{r}
ggplot(acc,aes(x=factor(HOUR),fill=factor(HOUR))) + geom_bar(stat_count='identity') + ylab('Number of Crashes Per Hour') + xlab("Hour of Day") + ggtitle("Number of Crashes Each Hour") + guides(fill=guide_legend("Days of the Month"))
```

###______Number of Crashes With Drunk Drivers______
```{r}
ggplot(acc,aes(x=factor(DRUNK_DR),fill=factor(DRUNK_DR))) + geom_bar(stat_count='identity') + ylab('Number of Crashes') + xlab("Number of Drunk Drivers") + ggtitle("Number of Crashes With Drunk Drivers")  + guides(fill=guide_legend("Number of Drunks"))
```

##Part Two:
#####Person Type 1 is the Driver of the vehicle
```{r}
person <- read.csv("https://raw.githubusercontent.com/xdaiISU/ds202materials/master/hwlabs/fars2017/person.csv", stringsAsFactors = FALSE)
filperson = filter(person, PER_TYP==1)
joined = inner_join(filperson, acc, by = "ST_CASE", type = "inner", match = "all")
```
###______Number of Crashes Each Day of the Week By Hour and Sex______
```{r}
joined = filter(joined, joined$HOUR.x < 25)
joined = filter(joined, joined$SEX < 3)
joined$DAY_WEEK = factor(joined$DAY_WEEK,labels=c("Sunday","Monday","Tuesday","Wednesday","Thursday","Friday","Saturday"))
ggplot(joined,aes(x=HOUR.x,fill=factor(SEX))) + geom_bar(position="dodge") + facet_wrap(joined$DAY_WEEK) + xlab("Hour of the Day") + ylab("Number of Crashes") + ggtitle("Number of Crashes Each Day of the Week by Hour and Sex") + guides(fill=guide_legend("SEX")) + scale_fill_discrete(name = "SEX", labels = c("Male","Female"))
```

##Part Three:
###______Number of Crashes Per County______
```{r}
GLC = readxl::read_xlsx("/Users/samuelshifflett/Documents/DS202/DS202_Github/ds202_lab5/ds202_lab5/GLC.xlsx")
colnames(GLC) = GLC[1, ] 
GLC = GLC[-1, ]
names(GLC)
counties = map_data("county")
names(GLC)[6] = "COUNTY"
names(GLC)[7] = "County_Name"
GLC$COUNTY = factor(GLC$COUNTY)
acc$COUNTY = factor(acc$COUNTY)
names(counties)[6] = "County_Name"
countacc = inner_join(acc,GLC, by="COUNTY") %>% select(LATITUDE,LONGITUD,COUNTY,"County_Name",DAY,MONTH,YEAR,HOUR,"State Name")
countacc = countacc %>% mutate(County_Name = tolower(County_Name))
countaccgroup = countacc %>% group_by(County_Name) %>% summarise(total_crashes = n())
countaccgroup = left_join(countaccgroup,counties, by="County_Name")
countaccgroup %>% group_by("County_Name") %>% ggplot(aes(x = long, y = lat,fill=total_crashes)) + geom_polygon(aes(group=group)) + ggtitle("Crashes Per County")  + guides(fill=guide_legend("Total Crashes")) + scale_fill_continuous(low="white",high = ("red"))
```

###______Ration of Crashes Per State by Season______
```{r}
states = map_data("state")
countaccseason = countacc %>% group_by(County_Name,MONTH) %>% summarise(total_crashes=n())
countaccseason = transform(countaccseason, SEASON = case_when((countaccseason$MONTH >= 11 | countaccseason$MONTH <= 3) ~ "WINTER", (countaccseason$MONTH >= 5 & countaccseason$MONTH <= 8) ~ "SUMMER"))
countaccseason = countaccseason[complete.cases(countaccseason[,4]),]
countaccseason = countaccseason %>% group_by(County_Name,SEASON) %>% summarise(total_crash = sum(total_crashes))  
countaccseason = countaccseason %>% left_join(counties,by="County_Name")
countaccseason = countaccseason %>% group_by(region,SEASON) %>% summarise(total_crashes=sum(total_crash))
countaccseason = countaccseason %>% spread(SEASON,total_crashes)
countaccseason = countaccseason %>% group_by(region) %>% summarise(ratioSUMbyWINT = SUMMER/WINTER)
countaccseason = countaccseason %>% left_join(states,by="region")
countaccseason %>% group_by("region") %>% ggplot(aes(x = long, y = lat,fill=ratioSUMbyWINT)) + geom_polygon(aes(group=group)) + ggtitle("Ratio of Total Crashes Per State (Summer/Winter)")  + guides(fill=guide_legend("Crash Percentage")) + scale_fill_continuous(low="red",high = ("white"))
```

As you can see the map is most red in Delaware meaning that the difference in crashes is the largest. Therefore Delaware has the worst time driving in the winter months.


=======
---
title: "team"
author: "cyber-atlas"
date: "April 11, 2019"
output: html_document
---

```{r}
library('ggplot2')
#acc <- read.csv("https://raw.githubusercontent.com/xdaiISU/ds202materials/master/hwlabs/fars2017/accident.csv", stringsAsFactors = FALSE)

acc<- read.csv('accident.csv', stringsAsFactors = FALSE)
class(acc)

#person <- read.csv("https://raw.githubusercontent.com/xdaiISU/ds202materials/master/hwlabs/fars2017/person.csv")

person$PER_TYP

person <- read.csv('person.csv', stringsAsFactors = FALSE)


```


```{r}
class(acc$DAY_WEEK)

acc$DAY_WEEK <- factor(acc[,'DAY_WEEK'])
levels(acc$DAY_WEEK) <- list("Sunday" = 1, 'Monday'=2, 'Tuesday'=3, 'Wednesday'=4, 'Thursday'=5, 'Friday'=6, 'Saturday'=7)

levels(acc$DAY_WEEK)

ggplot(data=acc, aes(x=DAY_WEEK, fill= DAY_WEEK))+
  geom_bar()+
  #geom_bar(aes(y=stat(count)))+
  geom_text(stat='count', aes(label=..count..), vjust = -.5)+
  scale_y_continuous(limits=c(0,6500))+
  theme(axis.text.x = element_text(angle =45, hjust=1))



```


Show the breakdown of the accidents by hour

```{r}
class(acc$HOUR)

acc$HOUR<- factor(acc[,'HOUR'])

levels(acc$HOUR)[levels(acc$HOUR) == 99] <- "unknown"

levels(acc$HOUR)

ggplot(data=acc, aes(x=HOUR, fill= stat(count), color=HOUR))+
         #fill=stat(count)))+
  geom_bar()+
  #geom_bar(aes(y=stat(count)))+
  geom_text(stat='count', aes(label=..count..), angle=90,hjust = -.5)+
  scale_y_continuous(limits=c(0,2500))+
  theme(axis.text.x = element_text(angle =75, hjust=1))
```



```{r}

class(acc$DRUNK_DR)

acc$DRUNK_DR<- factor(acc[,'DRUNK_DR'])

#levels(acc$DAY_WEEK) <- list("Sunday" = 1, 'Monday'=2, 'Tuesday'=3, 'Wednesday'=4, 'Thursday'=5, 'Friday'=6, 'Saturday'=7)

levels(acc$'DRUNK_DR')

ggplot(data=acc, aes(x=DRUNK_DR, fill=DRUNK_DR, color=stat(count)))+
         #fill=stat(count)))+
  geom_bar()+
 # geom_bar(aes(y=frequency(DRUNK_DR)))+
  geom_text(stat='count', aes(label=..count..), angle=0,vjust = -.1)
  #scale_y_continuous(limits=c(0,3000))
  #theme(axis.text.x = element_text(angle =75, hjust=1))
  #facet_wrap(~DRUNK_DR)


```


>>>>>>> fe5751fdab3423f170a6d9ec38ad5fb3ea16a849
